---
title: "Twitter Analysis of #OpenConCascadia19"
author: "Eric Leung"
date: "2019-02-04"
output:
  github_document:
    toc: true
    df_print: "kable"
---

# Objectives

# Data Management

```{r Look up tweets from ID, eval=FALSE}
# Saved file name
opencon_file <- "openconcascadia_search-ids.rds"
# Download status IDs file
download.file(
  "https://github.com/erictleung/openconcascadia19/blob/master/data/search-ids.rds?raw=true",
  opencon_file
)

# Read status IDs fromdownloaded file
ids <- readRDS(opencon_file)

# Lookup data associated with status ids
rt <- rtweet::lookup_tweets(ids$status_id)
```


# Packages!

```{r Load packages, warning=FALSE}
# Load main packages
library(rtweet)

# tidyverse and related packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(tidytext)

# Ancillary i.e., used once or twice in analysis
library(here)
library(syuzhet)
library(rvest)
```


# Query Tweets

```{r Query tweets with hashtag, message=FALSE}
# Search for "#OpenConCascadia19" hashtag
rt <- search_tweets(
  q = "#OpenConCascadia19",
  n = 2000,
  verbose = FALSE)
```

And let's save for easy dissemination.

```{r Save Twitter data}
# Create data directory if it doesn't exist
out_dir <- here("data")
dir.create(out_dir)

# Save out data
saveRDS(rt, here("data", "search.rds"))
saveRDS(rt, here("data", "search-ids.rds")) # Sharable data with just the IDs
```


# Explore Twitter

## Tweet frequency over time

The following summarizes Twitter data into a time series-like data frame. It is
then plotted over time (in 2-hour increments).

```{r Plot tweet frequency}
rt %>%
  filter(created_at > as.POSIXct("2019-01-31")) %>%
  ts_plot("2 hours", color = "transparent") +
  geom_smooth(method = "loess",
              se = FALSE,
              span = 0.1,
              size = 2,
              colour = "#0066aa") +
  geom_point(size = 5,
             shape = 21,
             fill = "#ADFF2F99",
             colour = "#000000dd") +
  theme_minimal(base_size = 15) +
  theme(axis.text = element_text(colour = "#222222"),
        plot.title = element_text(size = rel(1.2),
                                  face = "bold"),
        plot.subtitle = element_text(size = rel(0.9)),
        plot.caption = element_text(colour = "#444444")) +
  labs(title = "Frequency of tweets about #OpenConCascadia19 over time",
       subtitle = "Twitter status counts aggregated using two-hour intervals",
       caption = "\n\nSource: Data gathered via Twitter's standard `search/tweets` API using rtweet",
       x = NULL,
       y = NULL)
```


## Content sentiment as positive or negative

Let's do some sentiment analysis of our tweets. But before that, let's clea up
our tweets and remove some distractors from our analysis

```{r Clean tweets for sentiment}
# Clean up tweet text
rt <- rt %>%
  # Remove mentions
  # Rule are that names are alphanumeric and can have underscores.
  # Names can also be preceeded with "." or end with some punctuation
  # Twitter:
  #   https://help.twitter.com/en/managing-your-account/twitter-username-rules
  # To avoid emails:
  #   https://stackoverflow.com/questions/4424179/how-to-validate-a-twitter-username-using-regex#comment21201837_4424288
  mutate(clean_text = str_replace_all(text,
                                 "\\.?@([:alnum:]|_){1,15}(?![.A-Za-z])[:graph:]?",
                                 "")) %>%

  # Remove hashtags
  mutate(clean_text = str_replace_all(clean_text,
                                 "#[:graph:]+",
                                 "")) %>%

  # Remove links
  mutate(clean_text = str_replace_all(clean_text,
                                 "https?[:graph:]+",
                                 "'")) %>%

  # Convert to lowercase
  mutate(clean_text = str_to_lower(clean_text)) %>%

  # Trim extra write space
  mutate(clean_text = str_trim(clean_text)) %>%

  # Reduce repeated whitespace inside a string
  # E.g. str_squish("hello    world") -> "hello world"
  mutate(clean_text = str_squish(clean_text))
```

Now that we've cleaned up the tweets a bit we can do some sentiment analysis!

```{r Perform sentiment}
# Estimate sentiment for each tweet
rt <- rt %>%
  mutate(sentiment = get_sentiment(clean_text, "syuzhet"))

# Write function to round time into rounded var
round_time <- function(x, sec) {
  as.POSIXct(hms::hms(as.numeric(x) %/% sec * sec))
}

# Plot by specified time interval (1-hours)
rt %>%
  mutate(time = round_time(created_at, 60 * 60)) %>%
  group_by(time) %>%
  summarise(sentiment = mean(sentiment, na.rm = TRUE)) %>%
  mutate(valence = ifelse(sentiment > 0L, "Positive", "Negative")) %>%
  ggplot(aes(x = time, y = sentiment)) +
  geom_smooth(method = "loess",
              span = 0.1,
              colour = "#aa11aadd",
              fill = "#bbbbbb11") +
  geom_point(aes(fill = valence, colour = valence),
             shape = 21,
             alpha = 0.6,
             size = 3.5) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "none",
        axis.text = element_text(colour = "#222222"),
        plot.title = element_text(size = rel(1.2), face = "bold"),
        plot.subtitle = element_text(size = rel(0.9)),
        plot.caption = element_text(colour = "#444444")) +
  scale_fill_manual(values = c(Positive = "#2244ee", Negative = "#dd2222")) +
  scale_colour_manual(values = c(Positive = "#001155", Negative = "#550000")) +
  labs(x = NULL,
       y = NULL,
       title = "Sentiment (valence) of OpenCon Cascadia tweets over time",
       subtitle = "Mean sentiment of tweets aggregated in one-hour intervals",
       caption = "\nSource: Data gathered using rtweet. Sentiment analysis done using syuzhet")
```


## Popular emojis

Here is the full official list of emojis:
http://unicode.org/emoji/charts/full-emoji-list.html.

```{r Get all emojis}
emoji_page <- read_html("http://unicode.org/emoji/charts/full-emoji-list.html")
```

Now that we have these emojis, let's extract just the emoji stuff.

```{r}
emoji_tab <- emoji_page %>%
  html_node("table") %>%
  html_table(header = FALSE)
```

Almost there! Now just gotta fix up this table to be usable.

```{r}
emoji_check <- emoji_tab %>%
  select(X1, X2, X3, X15) %>%
  rename(ID = X1, code = X2, unicode = X3, shortname = X15) %>%
  filter(code != "Code") %>%
  filter(!str_detect(ID, "[:alpha:]")) %>%

  # Separate rows with various number of emojis
  mutate(num_spaces = str_count(code, " ")) %>%
  mutate(num_spaces_uni = str_count(unicode, " ")) %>%
  mutate(consistent = (num_spaces == num_spaces_uni))
  nest(-num_spaces) %>%
  mutate(data = map(data, function(df) {
    separate_rows(df, code, unicode)
  })) %>%
  unnest()
View(emoji_check)
```

Now that we have this table, we can do some analyses!

```{r Find popular emojis}

```


## Explore social network of Twitter attendees

```{r}

```


## Top associated hashtags

What were the most associated hashtags with OpenCon Cascadia?

```{r}

```


## How verbose are people tweeting?

```{r}

```


## Casing variation of hashtag

Which one was more popular? `#OpenConCascadia19`? Or `#openconcascadia19`?

```{r}

```

## Create chatterplot of common words

```{r}

```


# Commonly Shared Resources and Links

```{r}

```


# Helpful Resources in Analysis

Here are all of the links that have helped me analyze all of the above.

- Example Twitter analyses
  - [Data collection and visualization of #NCA17 tweets](https://github.com/mkearney/NCA17)
  - [Tracking and analyzing tweets about the 2018 National Communication Conference](https://github.com/mkearney/NCA18)
  - [A repository for tracking tweets about rstudio::conf](https://github.com/mkearney/rstudioconf_tweets)
- [Search emoji using `stringr`](https://stackoverflow.com/a/43370630) - I was
  going to try and remove emojis for my analysis, but I could find a way. This
  was the closest I could get to removing. It didn't work on my Windows machine.
- [Can We Please Stop Using Word Clouds](https://towardsdatascience.com/can-we-please-stop-using-word-clouds-eca2bbda7b9d) -
  Gave me the idea of using Sankey diagrams and then giving me the idea of using
  LDA from mentioning cluster maps.
- [RIP wordclouds, long live CHATTERPLOTS](https://towardsdatascience.com/rip-wordclouds-long-live-chatterplots-e76a76896098) -
  I found this alternative to word clouds really inspiring and useful.
- [Text Mining with R](https://www.tidytextmining.com) - General everything help
  on analyzing text, which I have no experience with.

# Session Information

```{r}
sessionInfo()
```
