---
title: "#NLMITC19 Twitter Analysis"
author: "Eric Leung"
output:
    md_document:
      toc: true
      df_print: "kable"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

## Load libraries

```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(ggrepel)

if (!requireNamespace("rtweet", quietly = TRUE)) install.packages("rtweet")
library(rtweet)
```


## Query data

Below is the code to query the Twitter data for the `#NLMITC19`. I ran this at
2019-06-28 22:50.

```{r query_tweets, eval=FALSE}
rt <- search_tweets("#NLMITC19 OR #NLMIT19", n = 1800, include_rts = FALSE)

saveRDS(rt, "nlmitc19_search.rds")
saveRDS(rt$status_id, "nlmitc19_search-ids.rds")
```

But instead, here I'll just look up the status IDs.

```{r read_in_data}
ids_file <- "nlmitc19_search-ids.rds"
nlmitc19_file <- "nlmitc19_search.rds"


# Read in search directly if exists
if (file.exists(nlmitc19_file)) {
  rt <- readRDS(nlmitc19_file)
} else {
  # Download status IDs file
  download.file(
    "https://github.com/erictleung/NLMITC19/blob/master/data/nlmitc19_search-ids.rds?raw=true",
    ids_file
  )

  # Read status IDs from downloaded file
  ids <- readRDS(ids_file)


  # Lookup data associated with status ids
  rt <- rtweet::lookup_tweets(ids)
}
```



## General tweet prevalence over time

Code modified from [`rstudioconf_tweets`][mk].

[mk]: https://github.com/mkearney/rstudioconf_tweets

```{r tweets_over_time, fig.height=7, fig.width=9}
rt %>%
  ts_plot("30 minutes", color = "transparent") +
  geom_smooth(method = "loess",
              se = FALSE,
              span = 0.05,
              size = 2,
              color = "#0066aa") +
  geom_point(size = 5,
             shape = 21,
             fill = "#ADFF2F99",
             color = "#000000dd") +

  # ggplot2 theme 
  theme_minimal(base_size = 15) +
  theme(axis.text = element_text(colour = "#222222"),
        plot.title = element_text(size = rel(1.7), face = "bold"),
        plot.subtitle = element_text(size = rel(1.3)),
        plot.caption = element_text(colour = "#444444")) +

  # Caption information
  labs(title = "Frequency of tweets about #NLMITC19 over time",
       subtitle = "Twitter status counts aggregated using half-hour intervals",
       caption = "\n\nSource: Data gathered via Twitter's standard `search/tweets` API using rtweet",
       x = NULL, y = NULL)
```

Makes sense considering there were two days of conference time.


## Most prolific tweeters?

```{r most_prolific_tweeter, fig.height=7, fig.width=9}
rt %>%
  group_by(screen_name) %>%
  summarise(tweets = n()) %>%
  ggplot(aes(x = tweets, y = reorder(screen_name, tweets))) +
  geom_point() +

  # Theme styling information
  theme_minimal(base_size = 15) +
  theme(axis.text = element_text(colour = "#222222"),
        plot.title = element_text(size = rel(1.7), face = "bold"),
        plot.subtitle = element_text(size = rel(1.3)),
        plot.caption = element_text(colour = "#444444")) +

  # Labels
  labs(title = "Top tweeters using\n#NLMITC19 or #NLMIT19",
       x = "Total number of tweets",
       y = "Twitter username",
       caption = "\n\nSource: Data gathered via Twitter's standard `search/tweets` API using rtweet")
```


## Relationship between follower count and tweet popularity

Do more followers have more popular tweets?

I take the average number of favorite of an individual's tweets and normalize it
based on the total number of tweets.

```{r follower_vs_favorites, fig.height=7, fig.width=9}
rt %>%
  # Preprocess and count average favorites normalized by number of tweets
  group_by(screen_name) %>%
  mutate(avg_fav = mean(favorite_count)) %>%
  mutate(avg_norm_fav = avg_fav / n()) %>%
  ungroup() %>%
  select(screen_name, avg_fav, avg_norm_fav, followers_count) %>%
  distinct() %>%

  # Offset to not create infinite values when log transforming
  mutate(followers_count = followers_count + 0.001) %>%
  mutate(avg_norm_fav = avg_norm_fav + 0.001) %>%

  # Plot results
  ggplot(aes(x = followers_count, y = avg_norm_fav, label = screen_name)) +
  geom_text_repel() +
  geom_point() +

  # Use log-scale for x-axis and y-axis
  labs(title = "Average normalized number of favorites\nversus user follower count",
       x = "Number of followers",
       y = "Average normalized number of favorites",
       caption = "\nSource: Data gathered via Twitter's standard `search/tweets` API using rtweet") +

  # Theme styling information
  theme_minimal(base_size = 15) +
  theme(axis.text = element_text(colour = "#222222"),
        plot.title = element_text(size = rel(1.7), face = "bold"),
        plot.subtitle = element_text(size = rel(1.3)),
        plot.caption = element_text(colour = "#444444"))
```


## Chatterplot of tweet words

```{r process_for_chatter}
rt_no_stop <- rt %>%
  # Just look at tweet text
  select(text, favorite_count) %>%
  
  # Remove web links
  mutate(text = str_replace_all(text, "https?[:graph:]+", "'")) %>%

  # Remove mentions
  # Rule are that names are alphanumeric and can have underscores.
  # Names can also be preceeded with "." or end with some punctuation
  # Twitter:
  #   help.twitter.com/en/managing-your-account/twitter-username-rules
  # To avoid emails:
  #   stackoverflow.com/questions/4424179/how-to-validate-a-twitter-username-using-regex#comment21201837_4424288
  mutate(text = str_replace_all(text,
                                "\\.?@([:alnum:]|_){1,15}(?![.A-Za-z])[:graph:]?",
                                "")) %>%

  # Tokenize text to just single words
  unnest_tokens(word, text) %>%

  # Remove stop words (e.g., "a", "the", "and", etc)
  anti_join(get_stopwords())


# Get average number of favorites
rt_word_avg_fav <- rt_no_stop %>%
  # Average favorite count
  group_by(word) %>%
  summarize(avg_fav = mean(favorite_count))


# Count number of mentions
rt_counts <- rt_no_stop %>%
  # Create word counts
  count(word, sort = TRUE)


# Filter low counts and join counts and average favorite score
chatter_rt <- rt_counts %>%
  filter(n > 1) %>%
  filter(word != "nlmitc19") %>%
  left_join(rt_word_avg_fav, by = "word")
```

Code below modified from ["RIP wordclouds, long live CHATTERPLOTS"][wordcloud].

[wordcloud]: https://towardsdatascience.com/rip-wordclouds-long-live-chatterplots-e76a76896098

```{r plot_chatter, fig.height=7, fig.width=9}
chatter_rt %>%
  # Add small offset average favorite counts because some are zero and we log
  # transform, which can introduce infinite values
  mutate(avg_fav = avg_fav + 0.001) %>%

  # Gather just top 100 mentions
  top_n(100, wt = n) %>%
  
  ggplot(aes(x = avg_fav, y = n, label = word)) +
  geom_text_repel(segment.alpha = 0,
                  aes(colour = avg_fav, size = n)) +

  # Set color gradient,log transform & customize legend
  scale_color_gradient(low = "green3", high = "violetred", 
                       trans = "log10",
                       guide = guide_colourbar(direction = "horizontal",
                                               title.position = "top")) +
  # Set word size range & turn off legend
  scale_size_continuous(range = c(3, 10),
                        guide = FALSE) +

  # Use log-scale for x-axis
  scale_x_log10() +
  ggtitle(paste0("Top 100 words from ",
                  nrow(rt),
                 " #NLMITC19 tweets, by frequency"),
          subtitle = "Word frequency (size) ~ Avg number of favorites (color)") + 
  labs(y = "Word frequency across all tweets",
       x = "Avg number of favorites in tweets containing word (log scale)",
       colour = "Avg num of favs (log)") +
  
  # minimal theme & customizations
  theme_minimal() +
  theme(legend.position = c(0.20, 0.99),
        legend.justification = c("right","top"),
        panel.grid.major = element_line(colour = "whitesmoke"))
```


## Session information

```{r}
sessionInfo()
```

